---
title: "case_study"
output: github_document
date: "2025-10-07"
---

```{r}
library(p8105.datasets)
library(tidyverse)
library(leaflet)
library(RColorBrewer)

data(nyc_airbnb)

```

## Brainstorm questions

*Pick a neighborhood / borough, investigate price
 * if we have dates, do that affect price?
*Does the type of unit affect rating / price?
*Does the name of the rental affect rating / price?
*Does rating vary by neighborhood, room type, or both?

*Where are rentals located?
*What rental has the highest review?
*How is price related to other variables?
*Are there any connections between price and other variables?
*What about prices over time?
 * Could we look at prices before and after the pandemic?
*How else did the pandemic affect AirBNB in NYC?
*What about ZIP codes?
*What does the number of days available have to do with anything else?


# neighborhood matters?

```{r}
nyc_airbnb |>
  drop_na(price) |>
  group_by(neighbourhood_group) |>
  summarize(
    mean_price = mean(price),
    min_price = min(price),
    max_price = max(price)
  )


```

## room type in a given neighborhood

```{r}
nyc_airbnb |>
  drop_na(price) |>
  group_by(room_type,neighbourhood_group) |>
  summarize(
    mean_price = mean(price)
  ) |>
  pivot_wider(
    names_from = room_type,
    values_from = mean_price
  )


```

```{r}
nyc_airbnb |>
  count(neighbourhood_group, room_type)

```




Make some plots

```{r}
nyc_airbnb |>
  ggplot(aes(x = price)) +
  geom_histogram()+
  scale_x_continuous(limits = c(0,1000))

```

```{r}
nyc_airbnb |>
  filter(
    neighbourhood_group == "Manhattan",
    price < 500,
    room_type == "Entire home/apt") |>
  mutate(neighbourhood = fct_recorder(neighbourhood, price)) |>
  ggplot(aes(x = neighborhood, y = price)) +
  geom_violin() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 900))


```

```{r}
nyc_airbnb |>
  filter(calculated_host_listings_count> 30) |> view()

```


Keep looking at location

```{r}
nyc_airbnb |>
  filter(
    neighbourhood_group == "Manhattan",
    price <1000,
    room_type == "Entire home/apt") |>
  ggplot(aes(x= lat, y = long, color = review_scores_location)) +
  geom_point()


```

a better map

```{r}
pal <-colorNumeric(
  palette = "viridis",
  domain = nyc_airbnb$review_scores_location)


nyc_airbnb |>
  filter(
    neighbourhood_group == "Manhattan",
    price <1000,
    room_type == "Entire home/apt") |>
  sample_n(25) |>
  leaflet() |>
  addTiles() |>
  addProviderTiles(providers$CartoDB.Positron) |>
  addCircleMarkers(
    ~lat, ~long,
    color = ~pal(review_scores_location),
    radius = 2)


```




# ZIP codes

```{r}

n_distinct(nyc_airbnb)




```






