---
title: "viz_1"
output: html_document
date: "2025-09-25"
---

```{r setup, include=FALSE}
library(ggplot2)
library(ggridges)
library(p8105.datasets)
library(tidyverse)

```

Import the weather data

```{r}
data("weather_df")

```

Make our first plot

```{r}

ggplot(data = weather_df, mapping = aes(x = tmin, y = tmax)) + 
  geom_point()


```


```{r}

ggp_weather_scatterplot = 
  weather_df |>
    ggplot(aes(x = tmin, y = tmax)) +
    geom_point()

ggp_weather_scatterplot


```

Check that some rows are missing

```{r}

weather_df |>
  filter(is.na(tmax))


```

## Fancier scatterplots

```{r}

weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.03) + 
  geom_smooth(se = FALSE)


```

Where you define aesthetics can matter

```{r}

weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name, size = prcp), alpha = 0.03) + 
  geom_smooth(se = FALSE)


```

use faceting real quick

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.3, size = 0.8) + 
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)


weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.3, size = 0.8) + 
  geom_smooth(se = FALSE) +
  facet_grid(name ~ .)


```

Let's make a somewhat more interesting 

```{r}

weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = 0.3) + 
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)


```

Learning Assessment: Write a code chain that starts with weather_df; focuses only on Central Park, converts temperatures to Fahrenheit, makes a scatterplot of min vs. max temperature, and overlays a linear regression line (using options in geom_smooth())


LA plot

```{r}

weather_df |> 
  filter(name == "CentralPark_NY") |> 
  mutate(
    tmax_fahr = tmax * (9 / 5) + 32,
    tmin_fahr = tmin * (9 / 5) + 32
    ) |> 
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
  geom_point(alpha = .5) + 
  geom_smooth(method = "lm", se = FALSE)
  


```

## small things

```{r}

weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) +
#  geom_point(aes(alpha = 0.3)) + 
  geom_smooth(se = FALSE)


weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) +
  geom_point() + 
  geom_smooth(se = FALSE)


```

```{r}

weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()


```

```{r}

weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(color = "purple")



```

## univariate plots

```{r}

weather_df |>
  ggplot(aes(x = tmin)) +
  geom_histogram(color = "white", fill = "red")
  
  
  
```



```{r}

weather_df |>
  ggplot(aes(x = tmin, color = name)) +
  geom_histogram()
  
weather_df |>
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram()


```

How would i fix this? maybe facet?

```{r}

weather_df |>
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram() +
  facet_grid(name ~ .)


```

a density plot?

```{r}


weather_df |>
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = 0.2)



```

```{r}

weather_df |>
  ggplot(aes(x = name, y = tmin)) +
  geom_boxplot(aes(fill = name))


```

ridge plot

```{r}

weather_df |>
  ggplot(aes(x = tmin, y = name, fill = name)) +
  geom_density_ridges()


```

LA univariate plots

```{r}

weather_df |>
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = 0.2)

weather_df |>
  ggplot(aes(x = prcp, fill = name)) +
  geom_histogram()+
  facet_grid(name ~ .)

weather_df |>
  filter(prcp>5)|>
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = 0.2)






```

## Saving and embedding plots

saving plots

```{r}

ggp_weather_violin = 
  weather_df |>
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin()

ggp_weather_violin

ggsave("plots/violin_plot.pdf", ggp_weather_violin,
       width = 8, height = 6)

```

embedding plots

```{r fih.width = 6, fig.height = 4}

ggp_weather_violin



```

